---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

Placeholder file for the future R coding exercise.

#Required packages and datasets uploaded here at the begining of the code
```{r}
library("dslabs")
library(ggplot2)
library(here)
```


#get discription, usage format and details fo the gapfinder dataset below
```{r}
help(gapminder)
```

#using the str command for a first look a the data
```{r}
str(gapminder)
```
#use the summary command to get some basic stats on the data 
```{r}
summary(gapminder)
```

```{r}
class(gapminder)
```

#select countries only in the  continient Africa, check number of observations and variables 

```{r}
africadata <- subset(gapminder, continent == "Africa")
dim(africadata)
```
# Examine structure of the new dataset created africadata
```{r}
str(africadata)
```
#run a summary command to see basic stats, number of observations for each
```{r}
summary(africadata)
```
#here create a subset of the africadata including only infant mortality and life expectancy
```{r}
africadata_mor<- africadata[,c("infant_mortality","life_expectancy")] 
```
#here create a subset of the afrcadata including only popuation and life expectancy
```{r}
africadata_ppln<- africadata[,c("population","life_expectancy")] 
```
#check structure of new dataset and number of observations of the two new datasets created africadata_mor and africadata_ppln using the comands str for strucutre and summary 
```{r}
str(africadata_mor)
```

```{r}
summary(africadata_mor)
```

```{r}
str(africadata_ppln)
```

```{r}
summary(africadata_ppln)
```

# plot life expectancy as a function of infant mortality as a function of population size.The geom commands includes the points on your graph.  Make two separate plots. Plot the data as points. For the plot with population size on the x-axis, set the x-axis to a log scale.

```{r}
ggplot(data=africadata_mor,aes(x=infant_mortality,y=life_expectancy) )+geom_point(lineend="butt", linejoin="round", linemitre=1)
```

#plot life expectancy as a function of population size, the scale_X_log10 command adds a third layer to the graph

```{r}
ggplot(data=africadata_ppln,aes(x=population,y=life_expectancy))+geom_point(lineend="butt", linejoin="round", linemitre=1) + scale_x_log10("population")

```

# to show years in the dataset  with missing data 

```{r}
africadata_filtered <-africadata[is.na( africadata$infant_mortality),]
africadata_unique <- unique(africadata_filtered$year)
```
#since the filtered dataset is an object, print it here to see the results showing the years with missing data for mortality 
```{r}
print(africadata_unique)
```
#create here a new object with a subset of the data that includes only the year 2000

```{r}
africadata_2000 <- subset(africadata, year == 2000)
dim(africadata_2000)
```

#check the expected observation numbers, stucture and summary of the new object created
```{r}
str(africadata_2000)
```

```{r}
summary(africadata_2000)
```

#plot to see correlations between infant mortality and life expectancy 
```{r}
ggplot(data=africadata_2000,aes(x=infant_mortality,y=life_expectancy) )+geom_point(lineend="butt", linejoin="round", linemitre=1)
```
#plot to see correlations between infant mortality and life expectancy 
```{r}
ggplot(data=africadata_2000,aes(x=population,y=life_expectancy))+geom_point(lineend="butt", linejoin="round", linemitre=1) + scale_x_log10("population")

```

#conduct a linear model life expectancy as outcome and infant mortality as the predictor. Use dataset subset created with the year 2000 only 

```{r}
fit1 <- lm(infant_mortality ~ life_expectancy, data=africadata_2000)
```
#check your model fit

```{r}
summary(fit1)
```

#conduct a linear model population as outcome and infant mortality as the predictor. Use dataset subset created with the year 2000 only 
```{r}
fit2 <- lm(infant_mortality ~ population, data=africadata_2000)
```

```{r}
#this chunk added by SETH Lattner
  #I believe that the lm() function calls for response~predictor (Y~X) so the models would be
fit1b<-lm(life_expectancy~infant_mortality, data = africadata_2000)
summary(fit1b)

  #and
fit2b<-lm(life_expectancy~population, data = africadata_2000)
summary(fit2b)

#in this case, model 1 would have significant results (p=2.826e-08)

```

#check your model fit
```{r}
summary(fit2)
```

#for your first model 
#use tidy broom function to save your model 

```{r}
fittable1 <- broom::tidy(fit1)
```
#check to see what your model looks like 
```{r}
print(fittable1)
```
#save your model as an object on your working directory using the here function 
```{r}
table_file1 = here("results", "resulttable1.rds")
saveRDS(fittable1, file = table_file1)
```

#for your second model 

```{r}
fittable2 <- broom::tidy(fit2)
```
#check to see what your model looks like 

```{r}
print(fittable2)
```
#save your model as an object on your working directory using the here function 

```{r}
table_file2 = here("results", "resulttable2.rds")
saveRDS(fittable2, file = table_file2)
```

Both model fits are not good. The predictor Infant mortality doesnt do a good job of explaining both popuation and life expectancy outcomes in this case.

### The following added by SETH LATTNER

For fun, lets do a bit more plotting on africadata, now looking at the gross domestic product (GDP). First, I want to calculate the per capita GDP.

```{r}

#calculate per capita gdp and add to new dataframe
gdp_per_cap<-africadata$gdp/africadata$population
summary(gdp_per_cap)
africadata_new<-cbind(africadata, gdp_per_cap)
```

Now, let's look at a few plots for each region of Africa

```{r, message=FALSE, warning=FALSE}
#boxplot of GDP per capita
ggplot(africadata_new, aes(region, gdp_per_cap))+
  labs(x="Region", y = "Per Capita GDP")+
  geom_boxplot(aes(fill=region), size= 0.5, alpha = 0.2)+
  theme_classic()+
  theme(legend.position = "none")


#plot infant mortality vs per capita GPD for each region from africadata
ggplot(data = africadata_new, aes(gdp_per_cap, infant_mortality, color=region))+
  geom_point(cex=2)+
  xlab("Per Capita GDP")+
  ylab("Infant Mortality (Per 1000 Individuals)")+
  theme_classic()
```