---
title: "Flu Data Analysis"
subtitle: "Fitting"
author: "Vijay Panthayir"
editor: visual
---

# Model Fitting

For the purpose of this exercise, the main predictor of interest is going to be `RunnyNose` and that we care about all other predictors. We are going to fit models with each of the two outcomes (body temperature and nausea) and `RunnyNose`. In addition, we are going to fit models with each of the two outcomes (body temperature and nausea) with all predictors. While fitting to one outcome, the other outcome will be considered a predictor (ex. when fitting predictors to body temperature, nausea will be considered a predictor).

This should all be completed with the `tidymodels` framework. We will be using the following commands:\
`linear_reg()`\
`set_engine("lm")`\
`logistic_reg()`\
`set_engine("glm")`

### 1. Load Data + Fitting Prep

First, lets load the packages needed for the fitting.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(recipes)
library(here)
library(parsnip)
library(dotwhisker)
library(performance)
```

Now, we can load and preview the processed data set.

```{r}
flu_clean <- readRDS(here::here("fluanalysis", "data","processed_data", "flu_data_processed"))
glimpse(flu_clean)
```

Finally, we can use `set_engine()` to create general linear and logistic models.

```{r}
linear <- linear_reg() %>% 
              set_engine("lm")
logistic <- logistic_reg() %>%
              set_engine("glm")
```

### 2. Linear Model Fitting

We are going to fit a linear model to the continuous outcome, body temperature, using the main predictor of interest, runny nose.

```{r}
lm_fit_rn <- linear %>%
              fit(BodyTemp~RunnyNose,data=flu_clean)
lm_fit_rn
```
### 3. Linear Model Fitting (cont.)

Now we are going to fit a linear model to the continuous outcome, body temperature, using all the predictors.

```{r}
lm_fit_all <- linear %>%
              fit(BodyTemp~ . ,data=flu_clean)
lm_fit_all
```

### 4. Linear Model Comparisons

```{r}
lm_fit_rn
lm_fit_all
compare_performance(lm_fit_rn,lm_fit_all)
```
### 5. Logistic Model Fitting

```{r}
glm_fit_rn <- logistic %>%
              fit(Nausea~RunnyNose,data=flu_clean)
glm_fit_rn
```

### 6. Logistic Model Fitting (cont.)

```{r}
glm_fit_all <- logistic %>%
              fit(Nausea~ . ,data=flu_clean)
glm_fit_all
```

### 7. Logistic Model Comparisons

```{r}
glm_fit_rn
glm_fit_all

compare_performance(glm_fit_rn,glm_fit_all)
```