---
title: "Flu Data Analysis"
subtitle: "Wrangling"
author: "Vijay Panthayi"
editor: visual
---

First, we load the tidyverse package.

```{r}
library(tidyverse)
```

Next, using the `here()` function and using the `readRDS` function, we import the `SympAct_Any_Pos.Rda` file.

```{r}
sympact_raw <- here::here("fluanalysis","data","raw_data","SympAct_Any_Pos.Rda")
flu_data_raw <- readRDS(sympact_raw)
summary(flu_data_raw)
```

The following are steps taken to clean the data set for exploration:

1.  We can use `!c()` with the `contains()` function to exclude variables that we do not want to analyze.
2.  Finally, we use `drop_na()` to remove all NA observations.
3.  We use `anyNA()` and `glimpse()` to determine if the previous steps worked properly.

```{r}
flu_data_clean <- flu_data_raw %>%
                select(!c(contains(c("Score","Total","FluA","FluB","Dxname","Activity")),"Unique.Visit")) %>%
                drop_na()
anyNA(flu_data_clean)
glimpse(flu_data_clean)
```

Finally, we save the cleaned data set into the `processed_data` folder using the `saveRDS()` function.

```{r}
flu_clean <- here::here("fluanalysis","data","processed_data","flu_data_processed")
saveRDS(flu_data_clean,file=flu_clean)
```
